---
title: "Diversity Indices"
output:
  bookdown::html_document2: 
    code_fold: hide
    number_sections: false
bibliography: references.bib
link-citations: yes
csl: apa-annotated-bibliography.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1) 
source("~/myWork/githubRepos/soenotes/R/shannonTest1.R")
source("~/myWork/githubRepos/soenotes/R/shannonTest2.R")
```

A note about making inferences on changes in direction of diversity indices, in particular the Shannon Index

$$Index = -\sum_{i = 1}^np_ilog(p_i)$$
where $n$ is the number of entities, $\sum p_i = 1$ and $p_i$ is the proportion of entity $i$'s contribution to the whole. This could be the proportion of species $i$'s biomass in the system, the proportion of total catch from species $i$, the proportion of fleet $i$'s revenue, effort etc.

## Properties of the index

The maximum value of the index is when when the system is evenly distributed, i.e. $p_i = 1/n$ for all $i$. This maximum increases as n increases

```{r maxes, echo = T,eval=T}
shannon <- NULL
nV <- 1:20
for(n in nV) {
  p <- rep(1/n,n)
  shannon[n] <- -sum(p*log(p))
}
plot(nV,shannon,type="l",xlab = "Number of Species",ylab = "Shannon Index")


```

## Simulation study

Can you make inferences based on the decline or increase in such an index over time if n remains unchanged? What about if n changes over time?

To investigate we need to simulate n probabilities conditional on their sum being equal to 1. We can do this in several ways, each way resulting in different distributional properties.

1. Simulate n uniform [0,1] random variables and divide by the sum (This will result in more even like probabilities)
2. Simulate n uniform random variables sequentially conditional on the previous value (This will result in more skewed, top heavy, distribution of probablites)
3. ?

### Eveneness

Let $X_i = U[0,1]$

Define $$p_i = \frac{X_i}{\sum_{j = 1}^nX_j}$$
Simulate 1000 samples for a range of n's.

```{r even, eval = T, echo = T}
p1 <- shannon_evenness(nSim = 1000,nspV=10:20)
plot(p1)
```

One can see that there is a significant amount of overlap between distributions making any conclusions about, lets say, a decline in the index complicated. It could be due to a reduction in the number of species or a reorganization of the entities in the system or a combination of both. And this is under the assumption that all entities are simulated somewhat evenly.

### Skewed

Let $X_1 = U[0,1]$
Then define:.

$X_2 = U[0,1-x_1]$

$X_3 = U[0,1-x_1-x_2]$ ...

$x_n = 1-\sum{x_i}$

Simulate 1000 samples for a range of n's

```{r skew, eval = T, echo = T}
p1 <- shannon_skewed(nSim = 1000,nspV=10:20)
plot(p1)
```

This random sampling is akin to having a few entities dominant in the system and all others insignificant. Under this sampling all distributions look similar. Very little inference could be made.

### Complete 

As can be seen by combining the above two methods of sampling there is a large part of the parameter space that is not being adequately sampled i.e the distribtuion of probabilities that result in values of the Shannon Index that exceed the value of 2 but are less than the lower tail of evenness.

Ideally we need a good method of simulating the full parameter space

## Conclusion

Inferences on the community structure, based on changes in this index, can not be made without deeper investigation into the composition of the entities that make up the system


## References

